Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-11-17 redaction/Presentations/SeminaireIrvine/slides_Irvine.tex

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\footnotesize
$$pq = \tau \widetilde{p}q$$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
\end{itemize}
\item[-] Some trade policy instruments as well (quota license price)
\footnotesize
$$pq = (t+\widetilde{p})q  $$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.2cm}
\begin{itemize}
\item[-] Many models rely on the tractability of iceberg costs \vspace{0.1cm}
\item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization because they distort relative prices (Sorensen, 2014)\vspace{0.1cm}
\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
\end{itemize}
\textbf{One objective of the paper is to provide an answer to this question}

@@

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.2cm}
\begin{itemize}
\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
%}
\end{itemize}
\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor linked to the importance of additive costs (which is kind of obvious) \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
%}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item[(3)] \textbf{ How have international transport costs evolved over time?} \vspace{0.2cm}
\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
\begin{itemize}
%\footnotesize{
\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
%}
\end{itemize}
\item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
\begin{itemize}
%\footnotesize{
\item[-] Not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2010) \vspace{0.2cm}
\item[-] Marked difference in the evolution of additive and multiplicative costs in Air
%\end{itemize}
%}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, large time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a larger period of time \vspace{0.1cm}
\item[-] Have both the import and the export prices: Estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 3-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
\end{itemize}

@@

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\vspace{0.1cm}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (3)}
\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
\begin{itemize}
\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs included:
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation}
%\normalsize
\begin{itemize}
\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
\end{itemize}
\item[-] With only ad-valorem costs:
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
%\normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item After running the estimates, we re-built: \vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[-] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

@@

 \begin{frame}
\frametitle{Result 1: Comments}
\begin{itemize}
\item Transport costs are sizeable \vspace{0.1cm}
\begin{itemize}
\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
\item[-] Lower than the 21\% markup taken from AVW, but  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
\end{itemize}
\end{frame}

\begin{frame}[label=slide_result2]
\frametitle{Result 2: Additive transport costs do matter}
\begin{itemize}
\item International transport costs: A sizeable additive component  \vspace{0.2cm}
\begin{itemize}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
\begin{itemize}
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.2cm}
\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.2cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport \vspace{0.2cm}
\end{itemize}
\item[-] A sizeable share of the additive component in total transport costs \vspace{0.2cm}
\begin{itemize}
\item[$\ast$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.2cm}
\item[$\ast$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_comment_compositioneffects]
\frametitle{Two main findings}
\begin{itemize}
\item The importance of excluding composition effects  \vspace{0.1cm}
\begin{itemize}
\item[-] The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
\item[-] Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
\end{itemize}
\item The importance of the additive component of transport costs (again) \vspace{0.1cm}
\begin{itemize}
\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in Air transport over the 1985-2005 period \vspace{0.1cm}
\item[-] In accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
\item[-] But... No substantial difference when additive costs are included (Panel (b))
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Decomposing \textit{pure} transport costs over time }
\begin{figure}[htbp]
%\caption{Decomposing transport costs over time,
%composition effects excluded}
%\label{fig:TC_add_mult_compeffects_excl}
\begin{center}
\includegraphics[width=8cm, height=3.5cm]{Fig3b_TCovertime_add_et_mult_3d.pdf}
%{\footnotesize {OECD data} }
\end{center}
\end{figure}
\begin{itemize}
\item For Vessel: Similar downward trend for both $\tau$ and $t$  \vspace{0.1cm}
\item For Air: Much more contrasted trends  \vspace{0.1cm}
\begin{itemize}
\item[-] Substantial decrease in the ad-valorem costs over 1985-2005, but roughly constant additive costs  \vspace{0.1cm}
\item[$\Rightarrow$] An explanation to the difference between Panels (a) and (b) above  \vspace{0.1cm}

\item[$\Rightarrow$] The importance of modeling additive costs  \vspace{0.1cm}
\item[-] Trend reversal around 2005: Any suggestion?
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Conclusion: Main findings}
\textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
\begin{itemize}
\item Provide a quantitative evaluation of both the additive and the ad-valorem components \vspace{0.1cm}
\begin{itemize}
\item[-] Based on the US imports flows from 1974 to 2013 \vspace{0.1cm}
\item[-] Additive cost: amount to 2.8\% of the export price in ocean shipping, 1.8\% in air transport \vspace{0.1cm}
\item[-] Iceberg cost: 3.2\% and 2.5\% for air and ocean respectively \vspace{0.1cm}
\end{itemize}
\item The importance of taking into account additive transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs are far from negligible quantitatively \vspace{0.1cm}
\item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
\end{itemize}
\item Characterize the evolution of transport costs over time \vspace{0.1cm}
\begin{itemize}
\item[-] Importance of the composition effects \vspace{0.1cm}
\item[-] Biased picture of the time trends of transport costs in air when omitting the additive dimension
\end{itemize}
\end{itemize}

>>>>

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\footnotesize
$$pq = \tau \widetilde{p}q$$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
\end{itemize}
\item[-] Some trade policy instruments as well (quota license price)
\footnotesize
$$pq = (t+\widetilde{p})q  $$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.2cm}
\begin{itemize}
\item[-] Many models rely on the tractability of iceberg costs \vspace{0.1cm}
\item[-] Additive trade costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization because they distort relative prices (Sorensen, 2014)\vspace{0.1cm}
\end{itemize}
\item Empiric studies
\begin{itemize}
\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2015, Hummels \& Skiba, 2004)  \vspace{0.1cm}
\item[-] Contrarian view  Lashkaripour (2016) taking into account the fact that more expensive goods are heavier and hence more expensive to transport. Yet  \vspace{0.1cm}
\begin{itemize}
\item[-] This can only apply to goods that are enumerated by items (60\% of US trade)
\item[-] Reasonable hypothesis for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods which importance has been rising since 1994 (the end point of Lashkaripour (2016))
\end{itemize}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
\end{itemize}
\textbf{One objective of the paper is to provide an answer to this question}

@@

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.2cm}
\begin{itemize}
\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
%}
\end{itemize}
\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
\item[$\Rightarrow$] We lose much. With the additive term included:\vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor linked to the importance of additive costs (which is kind of obvious) \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
%}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item[(3)] \textbf{ How have international transport costs evolved over time?} \vspace{0.2cm}
\item[$\Rightarrow$] The importance of excluding composition effects \vspace{0.2cm}
%\begin{itemize}
%\footnotesize{
%\item[-]\textit{Pure} transport costs have fallen since 1985 (not 1974), by $\simeq$ 40\% \vspace{0.2cm}
%}
%\end{itemize}
\item[$\Rightarrow$] The importance of including the additive component \vspace{0.2cm}
\begin{itemize}
%\footnotesize{
\item[-] Composition effect is especially strong for the additive costs \vspace{0.2cm}
\item[-] In the air case, the composition effec was not the same on additive and multiplicative costs
%\end{itemize}
%}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) (Could go up to HS-10 with recent data)\vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] The RHS are at the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 sectors (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 sectors at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
\item[-] We use both the import and the export prices. That allows the estimate the levels of both the ad-valorem and the additive trade costs ($\neq$ Irrarazabal et al., 2015, that uses one year of Norwegian data)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item Additional assumption : all products in a 3-digit sector (s) share the same structure of costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik^s}$ and $\tau_{ik^s}$ as:
\begin{equation}
 \tau_{ik^s} = \tau_{i} \times \tau_{s}, \qquad t_{ik} = t_{i} + t_{s} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 =\left(\tau_{i} \times \tau_{s} -1+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}} \right)\times \exp(\epsilon_{ik^s})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
\end{itemize}

@@

\begin{equation}
\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s} \label{eq:est_equation}
\end{equation}
\vspace{0.1cm}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not increase from one iteration to the next \vspace{0.1cm}
\item[-] Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (3)}
\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
\begin{itemize}
\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs included:
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}+\frac{t_{i} + t_{s}}{\widetilde{p}_{ik^s}}-1 \right) + \epsilon_{ik^s}
\end{equation}
%\normalsize
\begin{itemize}
\item[$\Rightarrow$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
\end{itemize}
\item[-] With only ad-valorem costs:
%\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik^s}}{\widetilde{p}_{ik^s}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s}-1 \right) + \epsilon^{ice}_{ik^s} \label{eq:model_nlI}
\end{equation}
%\normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item After running the estimates, we re-built: \vspace{0.1cm}
\begin{itemize}
\item[-] With additive costs:
$$\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}, \qquad \widehat{t}^{add}_{is} = \widehat{t}_{i} + \widehat{t}_{s}$$
\item[-] With only iceberg costs:
$$\widehat{\tau}^{ice}_{is} = \widehat{\tau_{i}} \times \widehat{\tau_{s}}$$
\end{itemize}
\item  Taking the weighted average over the sector-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

@@

 \begin{frame}
\frametitle{Result 1: Comments}
\begin{itemize}
\item Transport costs are sizeable \vspace{0.1cm}
\begin{itemize}
\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
\item[-] Lower than the 21\% markup taken from AVW, but  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] AVW decompose the 21\% markup in 9\% of time costs and 11\% in pure freight costs  \vspace{0.1cm}
\item[$\ast$] Their 11\% markup: A ``rough'' estimate, based on Hummels (2001) with 1994 data \vspace{0.1cm}
\item[$\ast$] Our data are only for the US, which presumably has lower trade costs \vspace{0.1cm}

\end{itemize}
\item[$\Rightarrow$] Our work: A more precise and exhaustive estimation of international transport costs \vspace{0.1cm}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
\end{itemize}
\end{frame}

\begin{frame}[label=slide_result2]
\frametitle{Result 2: Additive transport costs do matter}
\begin{itemize}
\item International transport costs: A sizeable additive component  \vspace{0.2cm}
\begin{itemize}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) (which is kind of obvios) \vspace{0.2cm}
\begin{itemize}
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.2cm}
\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.2cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport \vspace{0.2cm}
\end{itemize}
\item[-] A sizeable share of the additive component in total transport costs \vspace{0.2cm}
\begin{itemize}
\item[$\ast$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.2cm}
\item[$\ast$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_comment_compositioneffects]
\frametitle{What do we find}
\begin{itemize}
\item At the aggregate level  \vspace{0.1cm}
\begin{itemize}
\item[-] Air transport costs were reduced by 50\% and vessel transport costs by 60\%. (as already seen) \vspace{0.1cm}
\item[-] In the case of vessel, a large part of that is a composition effect (reduction in the share of bulky goods because of the relative decline of primary goods) \vspace{0.1cm}
\item[-] The reverse happened for air transport (increase in the shipping of bulky goods) \vspace{0.1cm}
\end{itemize}
\item Additive and multiplicative costs \vspace{0.1cm}
\begin{itemize}
\item[-] The composition effect is especially strong for the additive costs. Maybe bulkiness is more important for additive costs than for multiplicative costs ? \vspace{0.1cm}
\item[-] One might want to think of multiplicative as insurance and additive as shipping of handling. Will that help thinking about the composition effect? \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Conclusion: Main findings}
\textbf{Our paper: Empirical evidence about the role of the additive component in international transport costs} \vspace{0.1cm}
\begin{itemize}
\item Provide a quantitative evaluation of both the additive and the ad-valorem components \vspace{0.1cm}
\begin{itemize}
\item[-] Based on the US imports flows from 1974 to 2013 \vspace{0.1cm}
\item[-] Additive cost: amount to 2.8\% of the export price in ocean shipping, 1.8\% in air transport \vspace{0.1cm}
\item[-] Iceberg cost: 3.2\% and 2.5\% for air and ocean respectively \vspace{0.1cm}
\end{itemize}
\item The importance of taking into account additive transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs are far from negligible quantitatively \vspace{0.1cm}
\item[-] A better fit of the model when they are taken into account    \vspace{0.1cm}
\end{itemize}
\item Characterize the evolution of transport costs over time \vspace{0.1cm}
\begin{itemize}
\item[-] Importance of the composition effects \vspace{0.1cm}
\end{itemize}
\end{itemize}
